---
title: "Final Project"
author: "Andrew Zhuo"
date: '2022-07-06'
output: html_document
---

##What Are Chess Openings?

It is the initial stage of the chess game. Opening moves are based off chess theory, which are a set of general knowledges about analysis, positions, tactics and more.

##About the Project

Chess openings are crucial to a successful chess match for both sides. Are you ever curious about what types of openings grandmasters play the most frequently? This presentation will delve into the different and numerous openings 5 selected grandmasters play.


## Acquiring Data

We can first install the the necessary packages needed for this project.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(chessR)
library(tidyverse)
library(rvest)
library(dplyr)
library(plotly)
library(distill)
```



Now we can collect the data from the databases on chess.com. But the question at hand is which people should we use for the data. So we end up cho0sing the top 5 chess grandmasters currently: Magnus Carlsen, Ding Liren, Ian Nepomniachtchi, Alireza Firouzja and Fabiano Caruana.


```{r Load Data}
chess_openings <- get_raw_chessdotcom(usernames = c("MagnusCarlsen",
                                                    "Chefshouse",
                                                    "lachesisQ",
                                                    "Firouzja2003",
                                                    "FabianoCaruana"))
```

This is our current tibble we will be working with. Here is the raw data:

```{r Tibble}
glimpse(chess_openings)
```

##Processing Data

It's quite confusing, so we can search for the columns necessary: ECO (Encyclopedia of Chess Openings), the Username (for percentage calculations).

We can try finding the frequencies of each type of opening. Since ECO classifies various openings in different ways. A character in the beginning to represent the type of opening (i.e. flank opening, open game, etc.) So we can broadly classify these kind of openings by running this code snippet below (the string down there dictates the type of opening). So we can individually run the code and get the results:

-   18066 A type openings (flank)
-   10338 B type openings (semi-open games besides the French Defense)
-   5780 C type openings (open games)
-   3822 D type openings (semi-closed and closed games)
-   2585 E type openings (Indian Defenses)

Now we can proceed and create a plot, making good use of the data from the tibble:
```{r Most common type of opening}
graph_a <- chess_openings |> 
  select(ECO, Username) |> 
  mutate(total = n()) |> 
  drop_na() |> 
  #Get classification
  mutate(opening_type = ifelse(str_detect(ECO, "A"), "A", 
                              ifelse(str_detect(ECO, "B"), "B", 
                                     ifelse(str_detect(ECO, "C"), "C",
                                            ifelse(str_detect(ECO, "D"), "D", 
                                                   "E"))))) |> 
  mutate(games_played_with_opening = ifelse(opening_type == "A", 18066,
                                            ifelse(opening_type == "B", 10338, 
                                                   ifelse(opening_type == "C", 5780,
                                                          ifelse(opening_type == "D", 3822,
                                                                 2585))))) |> 
  mutate(percent = games_played_with_opening / 40591) |> 
  select(opening_type, percent) |> 
  distinct() |> 
  ggplot(mapping = aes(x = opening_type, y = percent, fill = opening_type)) + 
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  theme_classic() +
  labs(title = "Most Popular Type of Chess Openings Among 5  Grandmasters",
       subtitle = "'A' type (flank openings) appear to be the most commonly played",
       x = "Opening Type", y = "Occurrences", fill = "Type", 
       caption = "Source: chess.com databases")

ggplotly(graph_a, tooltip = c("x", "y", "color"))
```

It appears that Flank Openings (A) are most common among the grandmasters. Thus, to further investigate this, we use that data from column A and create a new graph this time analyzing the specific openings used.
```{r Most common opening}
y <- chess_openings |> 
  select(ECO, Username) |> 
  filter(str_detect(ECO, "A")) |> 
  #get percentage played
  mutate(games_for_opening = (n()/10000)) |> 
  arrange(ECO) |> 
  ggplot(mapping = aes(x = ECO, y = games_for_opening, fill = ECO)) +
  geom_col(width = 1.5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5)) +
  labs(title = "Comparing The Amount of Games Played Using Type-A Openings",
       subtitle = "A00 (Irregular openings) seems to be the most prevalent\n
       for some reason", x = "Opening", y = "Games played", 
       caption = "Source: chess.com databases")
ggplotly(y, tooltip = c("x", "color"))
```
##Summary
It appears that the most frequently used category of chess opening are flank openings, making up at least 44% of the data, hypermodern openings. But then again, the data is mostly from online games, which are usually more casual than actual games. Additionally, the selection of grandmasters for the sample might have been not diverse enough. Moreover, the library I used to compile the data might have had some issues with the data.

##Sources

- [Chess.com](chess.com)
- [RChess](https://github.com/jbkunst/rchess)
- And Primer.tutorials